---
title: "flights_feature_engineering"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Feature Engineering
For each flight:
_steer Change in compass bearing from the previous transponder detection for that aircraft; negative values indicate a turn to the left, positive values a turn to the right
_duration of each flight segment recorded by Flightradar24, in minutes.
_boxes Area of a rectangular bounding box drawn around each flight segment, in square kilometers

```{r message=FALSE}
# Load packages and set working directory
library(rjson)
library(dplyr)
library(tidyr)
library(ggmap)
library(ggplot2)
library(lubridate)
library(anytime)
library(sp)
library(tmaptools)
library(SpatialEpi)
library(Imap)

setwd("/Users/albertogonzalez/Documents/GitHub/mlearning/flights/eda_r")
list.files()

```

```{r message=FALSE}
# Read input data
input_file =read.csv(file = "flights_output_take_one.csv")
head(input_file)

```

# Initial calculations per one flight

```{r message=FALSE}
# Calculate duration: For each flight:
# duration of each flight segment recorded by Flightradar24, in minutes.
# flight_id = flight segment

# Select one flight
# Number of unique flights
u_flights = unique(input_file$adshex)
length(u_flights)
head(u_flights)

# Filter one and plot it (lat - lon)
one_flight = input_file %>%
  filter(adshex=="10534350")

str(one_flight)
head(one_flight)

# Flight segments per flight
segments = as.data.frame(table(one_flight$flight_id)) %>%
  arrange(desc(Freq)) %>% 
  head

segments

# Max and min timestamp
init = min(one_flight$timestamp)
init = anytime(init)

end = max(one_flight$timestamp)
end = anytime(end)
dura = difftime(end,init,units="mins")
dura

```
  
# Count of segments per aircraft

```{r message=FALSE}
segment_per_aircraft = input_file %>%
  group_by(adshex,flight_id) %>%
  summarise(n = n())

se = as.data.frame(segment_per_aircraft)

```



```{r message=FALSE}
# Calculate boxes: For each flight:
# boxes Area of a rectangular bounding box drawn around each flight segment, in square kilometers
bbox <- make_bbox(lon = one_flight$lng, lat = one_flight$lat, f = .1)
bbox

# Extract lon - lat
xy <- one_flight[,c(3,4)]

# Convert into spatial
# spdf <- SpatialPointsDataFrame(coords = xy, data = one_flight,
#                                proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))
# 
# class(spdf)
# 
# spat = SpatialPolygonsDataFrame(xy)
# 
# 
# # Calculate area in sq2
# area = approx_areas(spdf,target = "metric")

min_coord = head(xy,1)
max_coord = tail(xy,1)

coords = data.frame(rbind(min_coord,max_coord))
# area = latlong2grid(coords)
# area

area_meters = gdist(min_coord$lng,min_coord$lat,max_coord$lng,max_coord$lat, units = "m", a = 6378137.0, b = 6356752.3142, verbose = FALSE)

area_km2 = area_meters / 1000
area_km2

```



```{r message=FALSE}
# Calculate steer: change in compass from previous transponder
# Calculated over track
head(one_flight)
steer = diff(one_flight$track)
head(steer,10)

```


# Select 500 random flights to perform calculations


```{r message=FALSE}
head(input_file)
sample_df = sample_n(input_file, 500)
str(sample_df)

```

# Perform calculations for the 500 random selected flights

```{r message=FALSE}
# Duration
dura_df = sample_df %>%
  group_by(adshex) %>%
  summarize(min_time = min(timestamp),max_time = max(timestamp))

head(dura_df)


```


